CPU-Scheduling

### 2024-10-20 Handout/article.tex
Started cleaning up draft & added basic python structure

\section{Metrics}
MAYBE JUST LEAVING THIS OUT AND LOOKING AT THIS DURING A LATER TIME (WHEN THE METRICS ARE MENTIONNED OR LINK AS A FOOTNOTE?)
ALSO THIS FOLLOWS THE OSTEP WAY TOO MUCH:
We can measure different aspects of a Scheduler Policy. There is however a differentiation between fairness and performance. It is often a tradeoff between the two. For example the concept of everybody getting a same sized piece of cake is fair. However, it would be faster to give more to the fast eaters than to those who like to talk during eating. In this scenario we can't have both fairness and performance. In reality it is really difficult to predict, who'll eat faster.

@@

MAYBE HERE THE METRICS PART?

@@

\begin{lstlisting}[language=Python, style=colorEX]
# Simple implementation of FIFO in Python
list = []

# Adding a new process
def add_client(client):
  list.append(client)

@@

\begin{lstlisting}[language=Python, style=colorEX]
# Simple implementation of SJF in Python

# Define client class and list
class Client:
  def __init__(self, burst): # burst time is the time required for the process to finish
    self.burst = burst
    # other stuff that is relevant for the client class

@@

# Adding a new process
client = Client(burst) 
def add_client(client): 
  list.append(client)

@@

There are two notable changes in this code snippet, when compared to FIFO. One of them is relevant.
First, the less exiting part is that I actually took my time to write out some of the client class definition.
Second and the real difference to FIFO is that we sort the according to the burst time\footnote{Burst time is basically the runtime needed until the process finishes. This variable can be only predicted}. 
This means that we will choose the task with the shortest time to completion.
AT THIS PART THE KNOWLEDGE OF PROPER SORTING ALGORITHMS IS MISSING: MAYBE A REVERSE BUBBLE SORT? 
THE LIST SHOULD BE ALREADY SOMEWHAT ORDERED AND THE LAST MEMBER HAS TO BE SORTED
The only problem is when you arrive late. The scanning of products already started and you'll have to wait.
This policy is a so called non-preemptive policy.
"Non-preemptive" just means that once a task will run until it finishes.

@@

If SJF and FIFO are cousins that STCF is the brother.
Here we will make SJF "preemptive", meaning that our cashier can now "save" the state of the previously scanned items. 
Therefore STCF is also called "Preemptive Shortest Job First (PSJF)".
Although the switching costs some time, you'd have to wait anyways until our little boy returns with the ketchup\footnote{If you don't know what I am talking about read Chapter 1.1 (FIFO)}
The option of saving a state comes in handy, especially if a program has to wait for I/O.
The state of waiting is usually called "blocked" and the technique of running something else during the task is blocked is called overlapping.
Preemptiveness also makes the computer much more response, because now you can really do certain things "at the same time".
The best part is that we don't even have to change much of the previous code.
\pagebreak
\begin{lstlisting}[language=Python, style=colorEX]
  # Simple implementation of STCF in Python

  # definition of client class and list go here
  current = Client(burst)
  current_runtime = 0

  # Adding a new process
  def add_client(client): 
    list.append(client)
    allocate()

@@

    use_resource(current) # function that will allow "next" to use the CPU (also tracks runtime)
\end{lstlisting}
IDK ABOUT THE CODE; IT IS JUST KINDA HACKED TOGETHER; IT IS DEFINITELY NOT SECURE WHATSOEVER! 
Whenever a new client "joins", the "allocation" function gets run.
In addition to that, whenever the process finishes, or it voluntarily gives up the CPU, the same "allocation" function can be run.

@@

Now, why don't we just scan one item for everyone? 
This would mean that, even though slow but progress is being constantly made.
At this point the item scanning doesn't really work anymore\footnote{You can't scan a fraction of a product, or can you?} but let's just make the so called "quanta" (quantum in singular) even smaller.
As already mentionned in the previous chapter, the cashier needs some time to "save" the state of the current purchases. 
This effectively mean that after the quantum shrinks to a certain extent, the "context switching"\footnote{Context switch saves the state of a process so that it can be restored later} would get too expensive. 
The goal is to reach a balance, where we have enough responsibility, but the price for it is not too high.
Notice how by treating the processes more "fairly" by don't caring about their overall burst time, we get rid of a variable, which can only be predicted and replace it with another one, which is more abstract and global.
\pagebreak
\begin{lstlisting}[language=Python, style=colorEX]
  # Simple implementation of RR in Python

@@

  def add_client(client):
    list.append(client)

  def use_resource(client, quanta):
    # definition of use_resource
    # now it also takes an amount of quanta that the process can run for
    # after that it saves the state and returns.

  for client in list:
    use_resource(client, quantum)

@@

\noindent So as "OSTEP REFERENCE HERE" would summarize it: DO I EVEN NEED THIS?
\begin{enumerate}
  \item If Priority (A) $>$ Priority (B) $\Rightarrow$ A runs \& B doesn't
  \item If Priority (A) = Priority (B) $\Rightarrow$ A \& B run in RR
  \item When a job enters the system, it is placed at the highest priority
  \item Once a job uses up its time allotment at a given level, its priority is reduced
  \item After some period S, move all the jobs in the system to the topmost queue
\end{enumerate}

@@

Even though MLFQ introduces multiple Voo-Doo Constant, like the scheduling quantum, amount of queues, period S, allotment size (constant or changing?), we still gain an advantage over just simply unknown constant like the burst time or a predetermined priority and effort. These are huge improvements, because this means that we could actually implement it properly and the use case is not just purely a though experiment. 
It makes sense to find some sensible defaults for these constants and let the system admin adjust it, if they feel like it. One could find these magic numbers through machine learning, however that is out of scope of the topic?.
LINK THE ONE LINK HERE ABOUT ADJUSTING WITH MACHINE LEARNING

@@

For the whole script, please checkout my Github. There you can find

\begin{lstlisting}[language=Python, style=colorEX]
    # Implementation of MLFQ in Python

@@

\begin{itemize}
    \item \url{https://texblog.org/2015/09/30/fancy-boxes-for-theorem-lemma-and-proof-with-mdframed/}
    \item \url{https://wiki.osdev.org/Context_Switching}
    \item \url{https://en.wikipedia.org/wiki/Fernando_J._Corbató}
    \item \url{https://github.com/jkoritzinsky/xv6-MLFQ}
\end{itemize}

>>>>

\section{Metrics}
% COMMENT: MAYBE JUST LEAVING THIS OUT AND LOOKING AT THIS DURING A LATER TIME (WHEN THE METRICS ARE MENTIONNED OR LINK AS A FOOTNOTE?)
% COMMENT: ALSO THIS FOLLOWS THE OSTEP WAY TOO MUCH:
We can measure different aspects of a Scheduler Policy. There is however a differentiation between fairness and performance. It is often a tradeoff between the two. For example the concept of everybody getting a same sized piece of cake is fair. However, it would be faster to give more to the fast eaters than to those who like to talk during eating. In this scenario we can't have both fairness and performance. In reality it is really difficult to predict, who'll eat faster.

@@

% COMMENT: MAYBE HERE THE METRICS PART?

@@

\begin{lstlisting}[language=Python, style=colorEX, label=cd:fifo-python]
# Simple implementation of FIFO in Python
list = []

# Adding a new process
def add_process(process):
  list.append(process)

@@

\begin{lstlisting}[language=Python, style=colorEX, label=cd:sjf-python]
# Simple implementation of SJF in Python

# Define process class and list
class Process:
  def __init__(self, burst): # burst time is the time required for the process to finish
    self.burst = burst
    # other stuff that is relevant for the process class

@@

# Adding a new process
process = Process(burst) 
def add_process(process): 
  list.append(process)

@@

There are two notable changes in this code snippet, when compared to FIFO. One of them is relevant.
First, the less exiting part is that I actually took my time to write out some of the process class definition.
Second and the real difference to FIFO is that we sort the according to the burst time\footnote{Burst time is basically the runtime needed until the process finishes. This variable can be only predicted}. 
This means that we will choose the task with the shortest time to completion.
% COMMENT: AT THIS PART THE KNOWLEDGE OF PROPER SORTING ALGORITHMS IS MISSING: MAYBE A REVERSE BUBBLE SORT? 
% COMMENT: => Sorting algorithm input?
% COMMENT: THE LIST SHOULD BE ALREADY SOMEWHAT ORDERED AND THE LAST MEMBER HAS TO BE SORTED
The only problem is when you arrive late. The scanning of products already started and you'll have to wait.
This policy is a so called non-preemptive policy.
"Non-preemptive" just means that once a task will run until it finishes.

@@

If SJF and FIFO are cousins that STCF is the brother.
Here we will make SJF "preemptive", meaning that our cashier can now "save" the state of the previously scanned items. 
Therefore STCF is also called "Preemptive Shortest Job First (PSJF)".
Although the switching costs some time, you'd have to wait anyways until our little boy returns with the ketchup\footnote{If you don't know what I am talking about read Chapter 1.1 (FIFO)}
The option of saving a state comes in handy, especially if a program has to wait for I/O.
The state of waiting is usually called "blocked" and the technique of running something else during the task is blocked is called overlapping.
Preemptiveness also makes the computer much more response, because now you can really do certain things "at the same time".
The best part is that we don't even have to change much of the previous code.
\pagebreak
\begin{lstlisting}[language=Python, style=colorEX, label=cd:stcf-python]
  # Simple implementation of STCF in Python

  # definition of process class and list go here
  current = Process(burst)
  current_runtime = 0

  # Adding a new process
  def add_process(process): 
    list.append(process)
    allocate()

@@

    use_resource(current) # function that will allow "next" to use the CPU (also tracks runtime)
\end{lstlisting}
% COMMENT: IDK ABOUT THE CODE; IT IS JUST KINDA HACKED TOGETHER; IT IS DEFINITELY NOT SECURE WHATSOEVER! 
Whenever a new process "joins", the "allocation" function gets run.
In addition to that, whenever the process finishes, or it voluntarily gives up the CPU, the same "allocation" function can be run.

@@

Now, why don't we just scan one item for everyone? 
This would mean that, even though slow but progress is being constantly made.
At this point the item scanning doesn't really work anymore\footnote{You can't scan a fraction of a product, or can you?} but let's just make the so called "quanta" (quantum in singular) even smaller.
As already mentionned in the previous chapter, the cashier needs some time to "save" the state of the current purchases. 
This effectively mean that after the quantum shrinks to a certain extent, the "context switching"\footnote{Context switch saves the state of a process so that it can be restored later} would get too expensive. 
The goal is to reach a balance, where we have enough responsibility, but the price for it is not too high.
Notice how by treating the processes more "fairly" by don't caring about their overall burst time, we get rid of a variable, which can only be predicted and replace it with another one, which is more abstract and global.
\pagebreak
\begin{lstlisting}[language=Python, style=colorEX, label=cd:rr-python]
  # Simple implementation of RR in Python

@@

  def add_process(process):
    list.append(process)

  def use_resource(process, quanta):
    # definition of use_resource
    # now it also takes an amount of quanta that the process can run for
    # after that it saves the state and returns.

  for process in list:
    use_resource(process, quantum)

@@

\noindent So as "OSTEP REFERENCE HERE" would summarize it: % COMMENT: DO I EVEN NEED THIS?
\begin{enumerate}
  \item If Priority (A) $>$ Priority (B) $\Rightarrow$ A runs \& B doesn't
  \item If Priority (A) = Priority (B) $\Rightarrow$ A \& B run in RR
  \item When a job enters the system, it is placed at the highest priority
  \item Once a job uses up its time allotment at a given level, its priority is reduced
  \item After some period S, move all the jobs in the system to the topmost queue
\end{enumerate}

@@

Even though MLFQ introduces multiple Voo-Doo Constant, like the scheduling quantum, amount of queues, period S, allotment size (constant or changing?), we still gain an advantage over just simply unknown constant like the burst time or a predetermined priority and effort. These are huge improvements, because this means that we could actually implement it properly and the use case is not just purely a though experiment. 
It makes sense to find some sensible defaults for these constants and let the system admin adjust it, if they feel like it. One could find these magic numbers through machine learning, however that is out of scope of the topic?.
% COMMENT: LINK THE ONE LINK HERE ABOUT ADJUSTING WITH MACHINE LEARNING

@@

For the whole script, please checkout my Github.

\begin{lstlisting}[language=Python, style=colorEX, label=cd:mlfq-python]
  # Implementation of MLFQ in Python

@@

\begin{itemize}
    \item \url{https://texblog.org/2015/09/30/fancy-boxes-for-theorem-lemma-and-proof-with-mdframed/}
    \item \url{https://wiki.osdev.org/Context_Switching}
    \item \url{https://en.wikipedia.org/wiki/Fernando_J._Corbató}
    \item \url{https://github.com/jkoritzinsky/xv6-MLFQ}
    \item Obviously this is not complete and the main sources are not even here yet
\end{itemize}
